---
title: "Lab 19: Inference for Several Means"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(fig.height = 5)
knitr::opts_chunk$set(fig.width = 8.5)
knitr::opts_chunk$set(out.width = "100%")
knitr::opts_chunk$set(dpi = 300)

library(readr)
library(ggplot2)
library(dplyr)
library(viridis)
library(forcats)
library(smodels)
theme_set(theme_minimal())
```

## Instructions

Below you will find several empty R code scripts and answer prompts. Your task
is to fill in the required code snippets and answer the corresponding
questions.

## Burrito synergy in SoCal

Today we will look at a dataset of reviews of burritos in southern
California. This is a data set collected by a group of college friends who
live in the greater San Diego area.

```{r}
burrito <- read_csv("https://statsmaths.github.io/stat_data/burrito.csv")
```

The available variables in the data are:

- location: the name of the restaurant
- cost: total cost for the burrito
- yelp: the average Yelp review score
- google: the average Google review score
- chips_included: equals 1 if chips are included with the burrito
- hunger: score from 1 to 5 on how much the burrito filled up the reviewer
- tortilla: rating from reviewer; 1 (terrible) to 5 (awesome)
- temp: rating from reviewer; 1 (terrible) to 5 (awesome)
- meat: rating from reviewer; 1 (terrible) to 5 (awesome)
- fillings: rating from reviewer; 1 (terrible) to 5 (awesome)
- meat_filling: rating from reviewer; 1 (terrible) to 5 (awesome)
- uniformity: rating from reviewer; 1 (terrible) to 5 (awesome)
- salsa: rating from reviewer; 1 (terrible) to 5 (awesome)
- synergy: rating from reviewer; 1 (terrible) to 5 (awesome)
- wrap: rating from reviewer; 1 (terrible) to 5 (awesome)
- overall: rating from reviewer; 1 (terrible) to 5 (awesome)
- lat: latitude (in degrees North) of the taco restaurant
- lon: longitude (in degrees East) of the taco restaurant

Fit a linear regression predicting the average yelp score of locations in this
dataset and look at the model using `reg_table`:

```{r}

model_yelp <- lm_basic(yelp ~ 1, data = burrito)

reg_table(model_yelp, level = 0.95)

```


The average Yelp score of all restaurants in this area is just 3.7; looking
at the confidence interval from the model does it seem likely that burrito
locations have a higher Yelp rating than locations in general, or is this
just noise in the data?

**Answer**: It seems that the Yelp ratings would be higher in general. The lower bound of the confidence interval is 3.808, which is still higher than the average 3.7 Yelp ratings from all the restaurants in the area.

The regression model confidence interval requires that the data are sampling
independently from the larger population. Why might this dataset not hold to
this assumption?

**Answer**: Because the sampling came specifically from restaurants that sell burritos, so it could only be so random before it would have to be specified for the sample restaurants to sell burritos to be considered for the dataset. So it's not as random as it could have otherwise been.

Now fit a linear regression that predicts the yelp score as a function of
whether the burrito includes chips.

```{r}

model_yelp_chips <- lm_basic(yelp ~ 1 + chips_included, data = burrito)

reg_table(model_yelp_chips, level = 0.95)

```

Describe what the slope coefficient in the previous model really means
in terms of the data: 

**Answer**: This coefficient tells us that the average Yelp rating for restaurants with chips included is around 0.0569 higher than the average rating of restautants that do not include chips.

Is the slope coefficient statistically significant (compared to zero)?

**Answer**: No.

Describe this in terms of your answers above.

**Answer**: We could not confidently say that the average Yelp rating for restaurants including chips is significantly higher than the average ratings for the restuarants that don't include chips, because the lower bound is negative and the upper bound is positive. The confidence interval includes 0, which shows that it is still possible that there is no difference between the means.

The word *synergy* can be defined as:

> the interaction or cooperation of two or more organizations, substances,
> or other agents to produce a combined effect greater than the sum of
> their separate effects.

This is not something I would have thought of when considering tacos, but
let's see how synergy effected the scores of both the reviewers as well as
Google and Yelp users in general.

First, I'll add a variable called high_synergy defined as whether the synergy
score is greater than 3.5.

```{r}
burrito$high_synergy <- (burrito$synergy > 3.5)
```

Fit a regression predicting the Yelp score as a function of the
variable `high_synergy`.

```{r}

model_yelp_synergy <- lm_basic(yelp ~ 1 + high_synergy, data = burrito)

reg_table(model_yelp_synergy, level = 0.95)

```

Is there strong evidence that Yelp scores are higher for high synergy burrito
restaurants?

**Answer**: There is apparently no statistical significance for higher Yelp ratings for high synergy burrito restaurants.

Now fit a regression predicting the overall score as a function of the
variable high_synergy.

```{r}

model_overall <- lm_basic(overall ~ 1 + high_synergy, data = burrito)

reg_table(model_overall, level = 0.95)

```

Is there evidence that high synergy burrito locations have a higher overall
score than low synergy burrito locations?

**Answer**: Yes, there is evidence that the locations with high synergy burritos have statistically higher average overall scores than the locations with low synergy burritos.

Using the model from the previous question, what does the model predict
will be the overall score of a location with low synergy?

**Answer**: The score would be 2.696.

What does the model predict will be the overall score of a location with
high synergy?

**Answer**: The score would be 3.863.

Finally, fit a regression model that predicts the overall score as a function
of the variable `synergy` coded as a factor.

```{r}

model_overall_synergy <- lm_basic(overall ~ 1 + factor(synergy), data = burrito)

reg_table(model_overall_synergy, level = 0.95)

```

What does the model predict the overall score for a restaurant with a
synergy score of 3 will be?

**Answer**: It predicts the overall score for those restaurants will be 3.191.





