---
title: "Lab 17: Numeric Summaries"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(fig.height = 5)
knitr::opts_chunk$set(fig.width = 8.5)
knitr::opts_chunk$set(out.width = "100%")
knitr::opts_chunk$set(dpi = 300)

library(readr)
library(ggplot2)
library(dplyr)
library(viridis)
library(forcats)
library(smodels)
library(ggrepel)
theme_set(theme_minimal())
```

## Instructions

Below you will find several empty R code scripts and answer prompts. Your task
is to fill in the required code snippets and answer the corresponding
questions.

## Cereal Data

Today, we start by looking at a collection of breakfast cereals:

```{r}
cereal <- read_csv("https://statsmaths.github.io/stat_data/cereal.csv")
```

With variables:

- name: name of the specific cereal
- brand: name of the cereal's manufacturer
- sugar: amount of sugar per serving (g)
- score: healthiness score; 0-100; 100 is the best
- shelf: what shelf the cereal is typically stocked on in the store

Produce a histogram of the sugar variable.

```{r}

ggplot(cereal, aes(sugar)) +
  geom_histogram()

```

Now, compute the standard deviation of the variable `sugar`:

```{r}

sd(cereal$sugar)

```

What are the units of this measurement?


**Answer**: This shows the square root of how each brand varies in grams of sugar per serving from the mean.

Now, compute the deciles of the variable `score`:

```{r}

deciles(cereal$score)

```

What is the value of the 30th percentile. Describe what this means in words:

**Answer**: The value of the 30th percentile is 34.5. This means that about 30% of the cereal brands have a health score of less than 34.5 and that about 70% of the brands have a health score of greater than 34.5

Produce a boxplot of score and brand.

```{r}

ggplot(cereal, aes(brand, score)) +
  geom_boxplot()

```

Which brand seems to have the healthiest cereals?

**Answer**: Nabisco seems to have the healthiest cereals; about 50% of their cereals have a score above 60. 

Produce a boxplot of score and shelf.

```{r}

ggplot(cereal, aes(shelf, score)) +
  geom_boxplot()

```

Produce a boxplot of sugar and shelf.

```{r}

ggplot(cereal, aes(shelf, sugar)) +
  geom_boxplot()

```

If I want a healthy but reasonably sweet cereal which shelf would be the
best to look on?

**Answer**: I think you would look at the top shelf.

## Tea Reviews

Next, we will take another look at a dataset of tea reviews that I used in
a previous lecture:

```{r}
tea <- read_csv("https://statsmaths.github.io/stat_data/tea.csv")
```

With variables:
- name: the full name of the tea
- type: the type of tea. One of:
    - black
    - chai
    - decaf
    - flavors
    - green
    - herbal
    - masters
    - matcha
    - oolong
    - pu_erh
    - rooibos
    - white
- score: user rated score; from 0 to 100
- price: estimated price of one cup of tea
- num_reviews: total number of online reviews

Draw a scatterplot with num_reviews (x-axis) against score (y-axis) and add a
regression line (recall: `geom_smooth(method="lm")`).

```{r}

ggplot(tea, aes(num_reviews, score)) +
  geom_point() +
  geom_smooth(method = "lm")

```

Does the score tend to increase, decrease, or remain the same as the number
of reviews increases?

**Answer**: The score tends to increase as the number of reviews increases.

Calculate the ventiles of the variable price.

```{r}

ventiles(tea$price)

```

What is the 80th percentile? Describe it in words, include the units of the
problem in your answer.

**Answer**: The 80th percentile is 30.00. This means that about 80% of the teas cost less than 30.00 cents or less per cup, an that about 20% of the teas cost more than 30.00 cents per cup.

Plot the number of reviews (x-axis) against the score variable. Color
the points according to price binned into 5 buckets.

```{r}

ggplot(tea, aes(num_reviews, score)) +
  geom_point(aes(color = bin(price, 5)))
  

```

What tends to be true about the number of reviews for the most expensive
20% of teas?

**Answer**: The most expensive 20% of the teas tend to have the fewest number of reviews, relative to the other percentiles of teas.

Create a dataset named `white` that consists of only white teas.

```{r}

white <- filter(tea, type == "white")
white

```

Calculate the standard deviation of the price for white teas and the
standard deviation of the price for all of the teas.

```{r}

sd(white$price)
sd(tea$price)

```

Is the variation of the white tea prices smaller, larger, or about the same
as the entire dataset?

**Answer**: The variation in white tea prices is smaller than the variation in prices of all the tea types together.

Summarize the dataset by the type of tea and save the results as a variable
named `tea_type`.

```{r}

tea_type <- group_summarize(tea, type)

```

Plot the average price (x-axis) against the average score (y-axis) of
each type  of tea. Make the size of the points proportional to the number
of teas in each category and label the points with geom_text_repel and the
tea type.

```{r}

ggplot(tea_type, aes(price_mean, score_mean)) +
  geom_point(aes(size = n)) +
  geom_text_repel(aes(label = type))

```

Describe an interesting pattern or set of outliers that you found in the
previous plot. This does not need to take more than 1-2 sentences.

**Answer**: It's interesting that most of the tea types all fall into the same general area of the plot, being on average relatively cheap per cup but also having, on average, high ratings. I would think that perhps with cheaper teas, the mean scores would have been at risk for being lower, but most all of the cheaper teas were consistent on average with high ratings, even relative to the Masters brand.


