---
title: "Lab 22: Sixteen Personality Factors II"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(fig.height = 5)
knitr::opts_chunk$set(fig.width = 8.5)
knitr::opts_chunk$set(out.width = "100%")
knitr::opts_chunk$set(dpi = 300)

library(readr)
library(ggplot2)
library(dplyr)
library(viridis)
library(forcats)
library(smodels)
theme_set(theme_minimal())
```

## Instructions

Below you will find several empty R code scripts and answer prompts. Your task
is to fill in the required code snippets and answer the corresponding
questions.

# Title: "Sixteen Personality Factors II"


## 16PF Data, Again

Once again we will load the survey response data from the Sixteen Personality
Factor Questionnaire in order to practice our skills at statistical inference:

```{r}
pf <- read_csv("https://statsmaths.github.io/stat_data/cattell_16pf.csv")
```

The dataset use the following fields:
- age: respondent's age in years
- gender: respondent's self-selected gender
- country: two letter IATA code for the respondent's IP
- elapsed: time taken to complete quiz in seconds
- warmth: personality score from 1-20
- reasoning: personality score from 1-20
- emotional_stability: personality score from 1-20
- dominance: personality score from 1-20
- liveliness: personality score from 1-20
- rule_consciousness: personality score from 1-20
- social_boldness: personality score from 1-20
- sensitivity: personality score from 1-20
- vigilance: personality score from 1-20
- abstractedness: personality score from 1-20
- privateness: personality score from 1-20
- apprehension: personality score from 1-20
- openness_to_change: personality score from 1-20
- self_reliance: personality score from 1-20
- perfectionism: personality score from 1-20
- tension: personality score from 1-20
- baseline: average score across all 16 personality traits

Use the `percentiles` function to detect the 3th and 97th percentiles for
the amount of elapsed time taken to complete the quiz.

```{r}
percentiles(pf$elapsed)
```

**Answer**: The 3rd percentile is 369 seconds and the 97th percentile is 4107.46 seconds.

We want to construct a new variable called speed that tells us whether
someone was a fast test taker (less than the 3rd percentile), a slow test
taker (greater than the 97th percentile), or a normal test taker. To do
so, I'll use some code that we have not seen in this form before.
If the 3rd percentile was 100s and the 97th was 1000s, it would look like
this:

```{r}
pf$speed <- "normal"
pf$speed[pf$elapsed < 369.00] <- "fast"
pf$speed[pf$elapsed > 4107.46] <- "slow"
```

Modify the code above use the cutoffs you found in the previous question (and
make sure you run it).

Fit a linear model predicting the prefectionism variable as a function
of the speed variable.

```{r}

model <- lm_basic(perfectionism ~ 1 + speed, data = pf)

reg_table(model, level = 0.95)

```


How does the perfectionism score differ between the three groups? Are both
slopes significant?

**Answer**: The perfectionism score for the "fast" group is lower than the perfectionism score for the other two groups. The "normal" group increases at a higher rate than the "slow" group, and the slopes are all positively significant.

In the last question it should have appeared that the perfectionism
score for the slow and normal groups were very similar. In the regression
as given there is no way to test whether these two groups have a
statistically significant difference because the slopes only relate to the
baseline level (here, "fast").

It is possible to change the baseline level of a categorical variable in
R. To do so, use the function `fct_relevel`; its second argument gives the
desired baseline level:

```{r}

pf$speed_normal <- fct_relevel(pf$speed, "normal")

model <- lm_basic(perfectionism ~ 1 + speed_normal, data = pf)

reg_table(model, level = 0.95)

```

Using this table, is there a statistically significant difference between
normal and slow users in their perfectionism score?

**Answer**: There is no significant difference between the perfectionism scores for "normal" and "slow" users.

Based on the result in the previous describe a
plausible reason and a some people may finish quickly and a plausible reason
some may take a long time. Hint: Converting the 97th percentile to hours
may give you some ideas for the last part.

**Answer**: For the people who finished quickly--the "fast" group--who thus had a lower perfectionism score, they might have blown through the personality test at a faster rate because they didn't take the time to properly make sure they answered the questions in a way that most accurately described their personality. Therefore, their perfectionism scores would be lower because they didn't meticulously go through each item.
The perfectionism scores for the "slow" group were lower than those of the "normal" speed group, however the difference was not significant. Taking a bit over an hour to do the test, then, still yielded a high perfectionism score. The difference between "normal" and "slow" was not significant, so both still could qualify as having a relatively high perfectionism score that are not that different from each other.

## Selected trait, again

Take your personality trait from last class and select the one trait
that most closely contrasts with your trait and one that most closely
matches your trait. Fit a linear regression model that predicts your trait
as a function of both of these as well as the baseline variable.

```{r}

model <- lm_basic(warmth ~ 1 + privateness + liveliness + baseline, data = pf)

reg_table(model, level = 0.95)

```

Interpret the significance and signs of the slopes in the previous model
for the two traits that you selected. Do their signs match what you would
have expected?

**Answer**: The table's results do yield what I would have expected. The slope of privateness is significantly negatively related to warmth, and the slope for liveliness is significantly positively related to that of warmth. The trait I determined to be opposite of warmth (privateness) decreased significantly for every unit of increase in warmth, and the trait I determined to be most related to warmth (liveliness) increased significantly with every unit of increase for warmth.

Take the previous model and add the variables `speed`, `gender`, and
`country`, the latter lumped into 5 categories.

```{r}

model <- lm_basic(warmth ~ 1 + privateness + liveliness + baseline + speed + gender + fct_lump(country, n = 5), data = pf)

reg_table(model, level = 0.95)

```

Does this change the slopes for your two traits much?

**Answer**: The slopes did not change very much at all across warmth, privateness, and liveliness.

Summarize the previous model (specifically the relationship between the
three traits) in words.

**Answer**: This model shows the warmth scores on average as related to the baseline of Australia. Privateness scores were still significantly negatively related to warmth, and liveliness was still significantly positively related to warmth. The slope for India was the only country slope that was related to that of Australia, decreasing at a significantly negative rate. Females were still shown to have higher warmth ratings than males. The "speednormal" group still possessed higher warmth ratings than the other groups.

