---
title: "Lab 20: Linear Regression"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(fig.height = 5)
knitr::opts_chunk$set(fig.width = 8.5)
knitr::opts_chunk$set(out.width = "100%")
knitr::opts_chunk$set(dpi = 300)

library(readr)
library(ggplot2)
library(dplyr)
library(viridis)
library(forcats)
library(smodels)
theme_set(theme_minimal())
```

## Instructions

Below you will find several empty R code scripts and answer prompts. Your task
is to fill in the required code snippets and answer the corresponding
questions.

## Diamonds Data - Simple Regression

Today, we will start by looking at a dataset of diamond prices:

```{r}
diamonds <- read_csv("https://statsmaths.github.io/stat_data/diamonds.csv")
```

Fit a linear regression predicting the diamond price as a function of
its weight (carats). Print out the regression table without any confidence
intervals.

```{r}

model <- lm_basic(price ~ 1 + carat, data = diamonds)

reg_table(model)
  
```

How would you interpret the slope coefficient in this model?

**Answer**: As the price of the diamond increases, so too does the weight in carats.


Next, show the same regression as a scatter plot with a smoothing line:

```{r}

ggplot(diamonds, aes(price, carat)) +
  geom_point() +
  geom_smooth(method = "lm")

```

Now, print the same model as in the previous question but add a 95%
confidence interval.

```{r}

reg_table(model, level = 0.95)

```

What range of values does the confidence interval give for the intercept
in this model?

**Answer**: From -2282 to -2231.

What range of values does the confidence interval give for the slope
in this model?

**Answer**: From 7729 to 7784.

Would you be surprised if we collected another set of diamonds, in exactly the
same way as before, and found a slope estimate equal to 5500?

**Answer**: Yes, because that it outside of the confidence interval for the slope estimate.

Would you be surprised if we collected another set of diamonds, in exactly the
same way as before, and found a slope estimate equal to 7500?

**Answer**: Yes, because that too is outside of the confidence interval.

The biggest use of confidence intervals, at least for us, is to
determine whether the confidence interval for a slope contains only values
that are all negative, all positive, or a mix of both.

Why might this be important? Let's look at our diamond example. Describe the
meaning of the slope in this linear regression:

**Answer**: Because if the confidence interval is only positive or negative, that implies a significant relationship/correlation. So for this regression, the price would increase as the weight in carats increases, by around 7756.

Therefore, if the confidence interval has only positive values, we can
fairly confidently say is that there is a positive relationship between
diamond weight and its price.

Now, fit a regression model on the diamonds dataset, predicting the price as
a function of depth. Compute the regression table using level equal to .95:

```{r}

model_depth <- lm_basic(price ~ 1 + depth, data = diamonds)

reg_table(model_depth, level = 0.95)

```

What does the model predict will be the price of a diamond with a depth of
60?
```{r}
5763.67 + 60*-29.65
```


**Answer**: The price would be 3984.67.

Describe in words the estimate of the slope in terms of the dataset.

**Answer**: For every increase in the unit of depth of the diamonds, the price value will decrease by a value of -29.65.

Using the regression table from the previous question, does the
confidence interval contain only values that are positive, only values that
are negative, or values that are both positive and negative?

**Answer**: Only values that are negative, implying a significant negative relationship with price and depth.

Interpret this in words (your words, not just the formal definition).

**Answer**: So as the depth (in milimetres?) increases, the price of the diamond will drop by 29.65 for every milimetre increase in depth.

## Diamonds Data - Multiple Linear Regression

While we lose the nice graphical summary, it is possible to build linear
models with more than one variable and an intercept. In the diamond example,
for instance, we can fit a model that has both depth and carat in it:

```{r}
model <- lm_basic(price ~ 1 + carat + depth, data = diamonds)
reg_table(model, level = 0.95)
```

Notice that switching the order of the inputs does not change the estimates
or confidence intervals; only the order in the output is modified, but this
is inconsequential:

```{r}
model <- lm_basic(price ~ 1 + depth + carat, data = diamonds)
reg_table(model, level = 0.95)
```

Using the model in the previous question, what do you expect will be the
price of a diamond with a depth of 50 that weights 2 carats?

```{r}
4045.3 + -102.2*50 + 7765.1*2
```

**Answer**: The price would be 14465.5.

Are both of the slopes in the model statistically significant? What are
their signs?

**Answer**: Yes, they both are statistically significant. The sign for depth is negative and the sign for carat is positive.

Use the `add_prediction` function to add predictions from this model
back into the diamonds dataset.

```{r}

model <- lm_basic(price ~ 1 + depth + carat, data = diamonds)

diamonds <- add_prediction(diamonds, model)
diamonds

```

What is the largest positive residual (you can either use the `max` function
or sort the data in the data viewer)?

```{r}
max(diamonds$model_resid)
```

**Answer**: The largest positive residual is 12683.65.

Interpreting the meaning of the slopes in a regression model with
multiple variables changes slightly from the version with a single slope.
Each slope measures the change in the response give a change in its
corresponding value, when holding the other variables in the model fixed.
This last part is actually important and we will see some examples that at
first seem counterintuitive that make sense once we remember this caveat.

Describe the meaning the slopes, in your own words, in the regression model
above.

**Answer**: Seeing the negative slope for the depth, we can infer that for each increase in the unit of depth for the diamonds, the price will decrease at a rate of around 102.2. However, as the weight of the diamonds in carats increases, the price increases as well at a much higher rate, specifically by 7765.1. So although both slopes are significant, the slope of the carats has a stronger effect on the overall price than the depth of the diamonds.

Notice that we can do some comparisons quickly with the slope estimates.
For example, if we have two diamonds with the same depth where one is 1
carat in weight and the other is 2 carats in weight, how much more do you
expect the heavier diamond to cost?

```{r}
4045.3 + -102.2*50 + 7765.1*1
4045.3 + -102.2*50 + 7765.1*2

14465.5 - 6700.4
```


**Answer**: It would cost whatever the price of the less heavy diamond plus 7765.1 dollars. Since for every increase in carat, the price increases by 7765.1.


